Improving credibility and efficiency of decision-making \\ within democratic communities via a statistical quorum

### 2018-04-19 statistical-quorum.tex
crude related  work section done

\usepackage{todonotes}
\usepackage{bookmark}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{subcaption}

@@

\begin{enumerate}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed the novel concept of a \emph{statistical quorum} as a threshold of votes required to deem results credible and legitimate.
\item We implement a web application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in decision making processes.
\item We deployed the application in a democratically run community of 140
members for 16 weeks.
\item Using the analysis of decision making process before and after introduction
of the application, and using surveys,
%and using semi-structured interviews,
we show
how use of statistical quorum increases credibility of decisions,
improves utilization of collective attention, increases and diversifies overall participation,
and increases number of decisions made.
\end{enumerate}

The rest of the paper is structured as follows. In Section~\ref{sec:related-work} we
present related work. We continue by describing a framework of computing statistical
credibility of results and statistical quorum in Section~\ref{sec:decision-credibility}.
We present our methodology in Section~\ref{sec:methodology} after which we do
quantitative
%and qualitative
evaluation in Section~\ref{sec:evaluation}. We discuss results in
Section~\ref{sec:discussion} and conclude in Section~\ref{sec:conclusion}. At the end we
have Appendix where we derive formulas presented in Section~\ref{sec:decision-credibility}.

@@

@@

%Finally, we did semi-structured interviews after the fall semester with 7 members who were participating
%both in spring and fall semesters and ask them to compare their experience between those two semesters and
%evaluate the application and the use of statistical quorum.

@@

We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
Then we do quantitative
%and qualitative
evaluation through real-world use of the PeerMind application by a 140 members
community.

@@

Figure~\ref{fig:time} shows time spent per member on PeerMind, sorted by time. While on average, members
spent 7002 seconds (around 2 hours) using the application, the distribution of users is fairly
long-tailed with many members only spending 30 minutes and some members spending upwards of 20 hours.
On the other hand, based on minutes from spring semester we can observe that around 5 same members
regularly dominated conversation.

%\todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}

@@

%\subsection{Interviews}
%\label{sec:interviews}

%\todo[inline]{We made 7 unstructured interviews. Pull all quotes explaining how they like the idea and how it seems
%to work. But how there are other issues by moving decision making online: meetings also serve community building which
%does not happen online. No fun. So interesting insights about using online platforms for offline communities
%for making decisions which are not really answering our hypothesis, but are useful as insights we got. How some
%people prefer online and offline meetings.}

>>>>

\usepackage[disable]{todonotes}
\usepackage{bookmark}
\usepackage[T1]{fontenc}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{subcaption}

@@

\begin{enumerate}
\item We developed a framework of computing statistical credibility of results
based on sample size, votes cast, and voting duration.
\item Based on this framework we developed the novel concept of a \emph{statistical quorum} as a threshold of votes required to deem results credible and legitimate.
\item We implement a web application \emph{PeerMind},
integrating credibility of results and statistical quorum, and designed
a workflow of its use in decision making processes.
\item We deployed the application in a democratically run community of 140
members for 16 weeks.
\item Using the analysis of decision making process before and after introduction
of the application, using surveys, and using semi-structured interviews, we show
how use of statistical quorum increases credibility of decisions,
improves utilization of collective attention, increases and diversifies overall participation,
and increases number of decisions made.
\end{enumerate}

The rest of the paper is structured as follows. In Section~\ref{sec:related-work} we
present related work. We continue by describing a framework of computing statistical
credibility of results and statistical quorum in Section~\ref{sec:decision-credibility}.
We present our methodology in Section~\ref{sec:methodology} after which we do
quantitative and qualitative evaluation in Section~\ref{sec:evaluation}. We discuss results in
Section~\ref{sec:discussion} and conclude in Section~\ref{sec:conclusion}. At the end we
have Appendix where we derive formulas presented in Section~\ref{sec:decision-credibility}.

@@

Past approaches exist in using Bayesian statistics in the context of voting theory.  In particular, Bayesian methods have been proposed for auditing election outcomes \citep{rivest2012bayesian}.  Many further methods have been developed for efficient ballot auditing \cite{lindeman2012bravo}, \citep{rivest2017clipaudit}.  The problem of verifying an election's outcome is mathematically similar in structure to the notion of a statistical quorum defined in this work; however, a critical difference lies in the role the metric plays in the decision-making process itself.  Post hoc verification of results does not influence voting.  A statistical quorum, on the other hand, directly influences the allocation of a community's attention by essentially performing voting audits in real-time. To our knowledge, no methods have been formalized for coupling the verification of vote outcomes with the voting process itself. Furthermore, the dynamic statistical quorum presented here takes the previous ideas one step further.  Statistical method for auditing vote outcomes have no way to infer what we call the ``effective population size'', i.e., the portion of the population that actually cares about a vote's outcome.  We believe this information is crucial and have developed new statistical methods to infer this population and subsequently couple it to the statistical quorum, further influencing the influence on the decision-making process itself.  Additional work exists for selectively sampling a subset of the voting population to estimate the population's bias \cite{chaum2016random}. In random-sample voting, the emphasis is on security and efficiency of voting.  The approach taken to optimally allocate a community's attention is fundamentally different, however, in that the goal is to selectively distribute votes to random subsets of the population with the intent of reducing the amount of attention required by each individual.  This does indeed reduce the amount of attention required of individual voters, but at the same time is restricts the freedom of voters to participate only in the votes they have been delegated to participate in.

@@

Finally, we did semi-structured interviews after the fall semester with 7 members who were participating
both in spring and fall semesters and ask them to compare their experience between those two semesters and
evaluate the application and the use of statistical quorum.

@@

We evaluate statistical quorum in two ways. First, we analyze it theoretically and explore its mathematical properties.
Then we do quantitative and qualitative evaluation through real-world use of the PeerMind application by a 140 members
community.

@@

Figure~\ref{fig:time} shows time spent per member on PeerMind, sorted by time. While on average, members
spent 7002 seconds (around 2 hours) using the application, the distribution of users is fairly
long-tailed with many members only spending 30 minutes and some members spending upwards of 20 hours.

\todo[inline]{Using our metric compare legitimacy of decisions being made in the spring and fall.}

@@

\subsection{Interviews}
\label{sec:interviews}

\todo[inline]{We made 7 unstructured interviews. Pull all quotes explaining how they like the idea and how it seems
to work. But how there are other issues by moving decision making online: meetings also serve community building which
does not happen online. No fun. So interesting insights about using online platforms for offline communities
for making decisions which are not really answering our hypothesis, but are useful as insights we got. How some
people prefer online and offline meetings.}
